Embracing the Laws of Physics: \\ Three Reversible Models of Computation

### 2018-07-04 survey/survey.tex
wordcraft 4.4

We illustrate our methodology using a small example. Consider a
circuit that takes an input type consisting of three values
\Tree [ {\small a} [ {\small b} {\small c} ] ]~
and swaps the leftmost value with the rightmost value
to produce
\Tree [ {\small c} [ {\small b} {\small a} ] ]~.
We can implement two such circuits using our Agda library for $\Pi$:

@@

\noindent The first implementation rewrites the incoming values as follows:
\[
\Tree [ {\small a} [ {\small b} {\small c} ] ] ~\to~
\Tree [ [ {\small a} {\small b} ] {\small c} ] ~\to~
\Tree [ {\small c} [ {\small a} {\small b} ] ] ~\to~
\Tree [ {\small c} [ {\small b} {\small a} ] ] ~.
\]
\noindent
The second implementation rewrites the incoming values as follows:
\[
\Tree [ {\small a} [ {\small b} {\small c} ] ] ~\to~
\Tree [ {\small a} [ {\small c} {\small b} ] ] ~\to~
\Tree [ [ {\small a} {\small c} ] {\small b} ] ~\to~
\Tree [ [ {\small c} {\small a} ] {\small b} ] ~\to~
\Tree [ {\small c} [ {\small a} {\small b} ] ] ~\to~
\Tree [ {\small c} [ {\small b} {\small a} ] ] ~.
\]
\noindent The two circuits are extensionally equal. Using the level-2
isomorphisms we can \emph{explicitly} construct a sequence of
rewriting steps that transforms the second circuit to the first.  The proof
can be read as follows: the first three lines ``refocus'' from a right-associated
isomorphism onto the (left-associated) composition of the first $3$ isomorphisms;
then apply a complex rewrite on these (the ``hexagon'' coherence condition
of symmetric braided monoidal categories); this exposes two inverse combinators
next to each other --- so we have to refocus on these to eliminate them; we
finally re-associate to get the result.

>>>>

%%%%%%%%%
\subsection{Example}

We can now illustrate how this all works with a small example.
Consider a circuit that takes an input type consisting of three values
\Tree [ {\small a} [ {\small b} {\small c} ] ]~
and swaps the leftmost value with the rightmost value to produce
\Tree [ {\small c} [ {\small b} {\small a} ] ]~.
We can implement two such circuits using our Agda library for $\Pi$:

@@

\noindent The first implementation rewrites the incoming values as follows:
\[
\Tree [ {\small a} [ {\small b} {\small c} ] ] ~\to~
\Tree [ [ {\small a} {\small b} ] {\small c} ] ~\to~
\Tree [ {\small c} [ {\small a} {\small b} ] ] ~\to~
\Tree [ {\small c} [ {\small b} {\small a} ] ] ~.
\]
\noindent
The second implementation rewrites the incoming values as follows:
\[
\Tree [ {\small a} [ {\small b} {\small c} ] ] ~\to~
\Tree [ {\small a} [ {\small c} {\small b} ] ] ~\to~
\Tree [ [ {\small a} {\small c} ] {\small b} ] ~\to~
\Tree [ [ {\small c} {\small a} ] {\small b} ] ~\to~
\Tree [ {\small c} [ {\small a} {\small b} ] ] ~\to~
\Tree [ {\small c} [ {\small b} {\small a} ] ] ~.
\]
\noindent The two circuits are extensionally equal. Using the level-2
isomorphisms we can \emph{explicitly} construct a sequence of
rewriting steps that transforms the second circuit to the first.

We write such proofs in an equational style: in the left column, we have
the current combinator which is equivalent to the first one, and in
the right column, the justification for that equivalence. The
joining combinator is syntactic sugar for \transLR.  The transformation
could be written (using \transLR) by just giving all the pieces in
the right hand column --- but such transformations are very hard for
humans to understand and follow.

The proof
can be read as follows: the first three lines ``refocus'' from a right-associated
isomorphism onto the (left-associated) composition of the first $3$ isomorphisms;
then apply a complex rewrite on these (the ``hexagon'' coherence condition
of symmetric braided monoidal categories); this exposes two inverse combinators
next to each other --- so we have to refocus on these to eliminate them; we
finally re-associate to get the result.
